---
title: "Untitled"
author: "Group_13"
date: "2021/7/9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(moderndive)
library(infer)
library(broom)
library(knitr)
library(gridExtra)
library(GGally)
library(kableExtra)
library(corrplot)
library(RColorBrewer)
#library(olsrr)
library(car)
library(forcats)
library(ggfortify)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(olsrr)
```

# Interst Question
What influence do different features of coffee have on whether the quality of a batch of coffee is classified as good or poor?

# Data Discription
* country_of_origin – Country where the coffee bean originates from.  
* aroma – Aroma grade (ranging from 0-10)  
* flavor – Flavour grade (ranging from 0-10)  
* acidity – Acidity grade (ranging from 0-10)  
* category_two_defects – Count of category 2 type defects in the batch of coffee beans tested.  
* altitiude_mean_meters – Mean altitude of the growers farm (in metres)  
* harvested – Year the batch was harvested  
* Qualityclass – Quality score for the batch (Good - ≥82.5, Poor - <82.5). Note: 82.5 was selected as the cut off as this is the median score for all the batches tested.  

# Explantory Data Analysis

```{r data, echo = FALSE, eval = TRUE, warning = FALSE}
coffee <- read.csv("dataset13.csv",  fileEncoding = "UTF-8-BOM")
coffee <- coffee %>% mutate(Qualityclass = as.factor(Qualityclass), 
                            country_of_origin = as.factor(country_of_origin))

#change order for good and poor; take the poor as the response baseline
coffee$Qualityclass <- factor(coffee$Qualityclass, levels = c('Poor', 'Good'))

coffee.NoNa <- coffee %>% na.omit()
```

## Statistical summaries of the data

Let's show the numerical statistics of our data set.

```{r}
skim(coffee)
```
Missing values:
altitude_mean_meters: 201
harvested:             60

Outliers:
aroma, flavor, acidity, category_two_defects: p0 = zeros
category_two_defects:         p100 = 55
altitude_mean_meters: p0 = 1; p100 = 190164 (impossible high)


```{r}
#Guatemala: altitude_mean_meters 190164
#the highest elevation: 7492m 
#ref: https://en.wikipedia.org/wiki/List_of_elevation_extremes_by_country

#View the outliers altitude_mean_meters:190164, 190164, 110000, 11000
coffee.NoNa %>% top_n(n=10, wt = altitude_mean_meters) %>% arrange(desc(altitude_mean_meters))

coffee.No_NaOutliers <- coffee.NoNa %>% filter(altitude_mean_meters != 190164 & 
                                                altitude_mean_meters != 110000 &
                                                 altitude_mean_meters != 11000)


#Not quite sure the altitude_mean_meters == 1, should be dropped
coffee.No_NaOutliers %>% select(country_of_origin, altitude_mean_meters) %>% arrange(altitude_mean_meters)

#View the outlier in aroma, flavor, acidity = 0 is Honduras (cause problem in logistic)
coffee.No_NaOutliers %>% filter(aroma == 0)
coffee.No_NaOutliers_keepHonduras <- coffee.No_NaOutliers
coffee.No_NaOutliers <- coffee.No_NaOutliers %>% filter(aroma != 0)


skim(coffee.No_NaOutliers_keepHonduras)
skim(coffee.No_NaOutliers)

```

#Graph

```{r cor}
ggpairs(coffee.No_NaOutliers[c(-1,-7)], aes(alpha = 0.4))
```
flavor & aroma   = 0.725
flavor & acidity = 0.744
aroma  & acidity = 0.591
an outlier?

```{r boxplot}
coffee.No_NaOutliers %>% ggplot(aes(x = country_of_origin, y = altitude_mean_meters))+
  geom_boxplot()

coffee.NoNa %>% ggplot(aes(x = country_of_origin, y = category_two_defects))+
  geom_boxplot()
```

```{r boxplot by Qualityclass}
p1 <- coffee.No_NaOutliers %>% ggplot(aes(x = Qualityclass, y= aroma)) + 
  geom_boxplot()

p2 <- coffee.No_NaOutliers %>% ggplot(aes(x = Qualityclass, y= flavor)) + 
  geom_boxplot()

p3 <- coffee.No_NaOutliers %>% ggplot(aes(x = Qualityclass, y= acidity)) + 
  geom_boxplot()

p4 <- coffee.No_NaOutliers %>% ggplot(aes(x = Qualityclass, y= category_two_defects)) + 
  geom_boxplot()

p5 <- coffee.No_NaOutliers %>% ggplot(aes(x = Qualityclass, y= altitude_mean_meters)) + 
  geom_boxplot()

grid.arrange(p1,p2,p3,p4,p5)
```


```{r m1}
m1 <- glm(Qualityclass ~ ., family = binomial(link = "logit"), data = coffee.No_NaOutliers[,c(-1, -7)]) #no years & country

summary(m1) #AIC 562.26
summ(m1)
glance(m1)$AIC
```

```{r vif}
vif(m1)
```
VIF check: they all below 10


```{r plot_m1_log-odds}
plot_model(m1, show.values = TRUE, transform = NULL,
           title = "Log-Odds", show.p = FALSE)
```

```{r m2 no_altitude_mean_meters}
m2 <- glm(Qualityclass ~ aroma + flavor + acidity + category_two_defects, family = binomial(link = "logit"), data = coffee.No_NaOutliers[,c(-1, -7)]) #no years & country

summary(m2) #566.59
glance(m2)$AIC
```

```{r m3_no_category_two_defects}
m3 <- glm(Qualityclass ~ aroma + flavor + acidity + altitude_mean_meters, family = binomial(link = "logit"), data = coffee.No_NaOutliers[,c(-1, -7)]) #no years & country

summary(m3) #560.26 (the best one)
#The best one is m3, explanatory variables: flavor, aroma, acidity,  altitude_mean_meters
glance(m3)$AIC
```
the positive coef for flavor covariate suggests that a higher chance of good quality for higher grade of flavor.


```{r m4 aroma, flavor, acidity}
m4 <- glm(Qualityclass ~ aroma + flavor + acidity, family = binomial(link = "logit"), data = coffee.No_NaOutliers[,c(-1, -7)]) #no years & country

summary(m4) #564.62
glance(m1)$AIC
```

```{r stepwise}
full.m1 <- m1
full.m2 <- glm(Qualityclass ~ ., family = binomial(link = "logit"), data = coffee.No_NaOutliers[,c(-1)]) #include harvested(year) as explanatory variable

#The same as m3 (full.m1,full.m2 are the same)
both.frmFull<-step(full.m2, scope = list(upper = full.m2),direction = 'both') 

#The same as m3 (full.m1,full.m2 are the same)
null <- glm(Qualityclass ~ 1, family = binomial(link = "logit"), data = coffee.No_NaOutliers[,c(-1)])
both.frmNull<-step(null, scope = list(upper = full.m2),direction = 'both') 
```
The best one is m3, explanatory variables: flavor, aroma, acidity,  altitude_mean_meters

```{r deviance check}
m3$null.deviance - m3$deviance > qchisq(df = m3$df.null-m3$df.residual, 0.95)
#APM page 19 (compare null model)
#True = reject null hypothesis
```
We can reject null hypothesis, and conclude that flavor, aroma, acidity, altitude_mean_meters are significant term in the model.

```{r anova for m3}
anova(m3)
qchisq(df = 1, p = 0.95)
```
It can be perceived that the largest reduction in residual deviance comes from aroma(472.06) and the smallest reduction in residual deviance is altitude_mean_meters(6.35). Even though altitude_mean_meters contributes the smallest reduction of residual deviance, this is larger than the 95th percentile of $X^2(1)$ distribution 3.841459. Therefore, this does not need to drop.

```{r comp, echo = FALSE, eval = TRUE, warning = FALSE}
model.comp.values.mlr.model1 <- glance(m1)
model.comp.values.mlr.model2 <- glance(m2)
model.comp.values.mlr.model3 <- glance(m3)
model.comp.values.mlr.model4 <- glance(m4)

Models <- c("MLR1","MLR2","MLR3", "MLR4")
bind_rows(model.comp.values.mlr.model1, model.comp.values.mlr.model2,model.comp.values.mlr.model3, model.comp.values.mlr.model4, .id="Model") %>%
  select(Model, null.deviance, deviance,AIC,BIC) %>%
  mutate(Model=Models) %>%
  kable(
    digits = 2, align = "c", col.names = c("model","null.deviance","deviance","AIC","BIC"),
    caption = "\\label{tab:select} Model comparison values for different models",
  )%>%
  kable_styling(font_size=10, latex_options="HOLD_position")
```


```{r Hosmer-Lemeshow goodness of fit test}
source(url("http://www.chrisbilder.com/categorical/Chapter5/AllGOFTests.R"))
HLTest(m3, g =10)
HLTest(m3, g =6)
```
When fitted value group into 10 classes, p-value = 3.638e-09
When fitted value group into 6 classes, p-value = 0.002889
The p-value is not large; this indicates that the m3 is not a good fit for the data (observed frequencies are far from expected frequencies)
However, some expected counts are less than 5; this suggests that the chi-squared approximation my not be very reliable for these data



```{r plot_m3_log-odds}
plot_model(m3, show.values = TRUE, transform = NULL,
           title = "Log-Odds", show.p = FALSE)
```

```{r  plot_m3_odds}
plot_model(m3, show.values = TRUE,
           title = "Odds", show.p = FALSE)
```


```{r plot_m3 glm repectively}
plot_model(m3, type = "pred", title = "")
```


```{r glm coffee.No_NaOutliers_keepHonduras}
#problem over-fitted since  
glm(Qualityclass ~ ., family = binomial(link = "logit"), data = coffee.No_NaOutliers_keepHonduras[,c(-1, -7)])
```



```{r}
library(glmnet)
autoplot(m3)
```

